Estimating Correspondences of Deformable Objects ``In-the-wild"

### 2016-04-07 contents/implementation.tex
Updates from ShareLaTeX
Final
minor changes and acknowledgments

%\subsection{Correspondence Basis for Shape Flow Estimation}
{\refstepcounter{steps}\label{sec:step2}\subsection*{Step 2: Correspondence Basis for Shape Flow Estimation}}

@@

Using the constructed correspondence basis, the shape flow can be approximated as:
\begin{equation}\label{eq:LinTrajModel}
    \bm{u}_n(\bx) \approx
    \textstyle \, \sum_{i=1}^R \, \bm{q}_i(n)\bm{v}_i(\bx) \,\, ,
\end{equation}
where $\bm{v}_i(\bx)$ is the weight that needs to be applied on the $i$-th correspondence basis element, in order to get the correspondence vector for the point location $\bm{x}$. In other words, the shape flow for every point $\bm{x}$ is described as a linear combination of basis elements that is controlled by the coefficients $\bm{v}_i(\bm{x})$.
The values of the $i$-th coefficient for all the points $\bm{v}_i(\bm{x})$ can be interpreted as an image defined on $\Omega$. Using the correspondence basis, the determination of the shape flow boils down to the determination of the set of coefficients $\bm{v}_i(\bm{x})$. The above representation of shape flow, constrains the correspondence vectors to lie on a subspace and, therefore, acts as a low-rank prior that enforces coherency of the shape registration result over the whole training dataset of shapes.

@@

\begin{figure}[b!]
    \centering
    \includegraphics[width=0.4\textwidth]{resources/models}
    \caption{Exemplar dense shape models build for faces and ears. Dense shapes are presented as grid for better visualisation.}
    \label{fig:dense_models}
\end{figure}

\begin{figure*}[t!]
\begin{center}
 \begin{tabular}{ccc}
\subfigure[]{
    \newcommand{\flowh}{0.12\columnwidth}
    \includegraphics[height=\flowh]{resources/Fig_Flows/1}
    \includegraphics[height=\flowh]{resources/Fig_Flows/2}
    \includegraphics[height=\flowh]{resources/Fig_Flows/3}
    \includegraphics[height=\flowh]{resources/Fig_Flows/4}
    \includegraphics[height=\flowh]{resources/Fig_Flows/5}
    \includegraphics[height=\flowh]{resources/Fig_Flows/6}
} &
\subfigure[]{
    \includegraphics[width=0.2\textwidth]{resources/models}
}
 \end{tabular}
\end{center}
\caption{(a) . (b) }
\label{REC}
\end{figure*}

% ---------------------------------------------------------------------------------------------------------------------------------------------------

{\refstepcounter{steps}\label{sec:step3}\subsection*{Step 3: Shape Flow Estimation}}
% \subsection{Shape Flow Estimation}
% \label{sec:step3}

@@

Let $\bm{d}(\bx;n)$, $\bm{d}(\bx;0):\Omega\rightarrow \R^{N_c}$ be the $n$-th training SVS image and the reference SVS image respectively. Following \cite{Garg:2013hu}, we propose to estimate the shape flow over all training images by minimizing the following energy:
\begin{align}
E_{sf} & =\alpha
\int_{\Omega}\sum_{n=1}^{N_t} \|\bm{d}(\bx+\bm{u}_n(x);n)-\bm{d}(x;0)\| \ud \bx \label{eq:costfunc}\\
    &+ \beta \int_{\Omega}\sum_{n=1}^{N_t}\|\bm{u}_n(\bx)-\sum_{i=1}^R\bm{q}_i(n)\bm{v}_i(\bx)\|^2 \ud \bx \label{eq:lowrank}\\
    &+
\int_\Omega  \sum_{i=1}^R \,\, \left \|    \nabla \bm{v}_i(\bx)    \right \|  \,\ud \bx \label{eq:TVterm}
\end{align}
This energy consists of two sets of unknown shape flows that are relatively close to each other: (i) $\bm{u}_n(x)$ which tries to explain the data from the input SVS images, and (ii) the shape flow determined by the correspondence basis coefficients $\bm{v}_i(\bx)$ that are spatially regularised and enforce a low-rank prior. 
%We minimize this energy jointly with respect to $\bm{u}_n(\bx)$ and $\bm{v}_i(\bx)$. 
The positive parameters $\alpha$ and $\beta$ weigh the balance between the terms of the energy.

@@

{\refstepcounter{steps}\label{sec:step4}\subsection*{Step 4: Dense and Patch-Based Deformable Models}}
% \subsection{Dense and Patch-Based Active Appearance Models}

>>>>

\vspace{0.2cm}
%\subsection{Correspondence Basis for Shape Flow Estimation}
{\refstepcounter{steps}\label{sec:step2}\subsection*{Step 2: Correspondence Basis for Shape Flow Estimation}}

@@

Using the constructed correspondence basis, the shape flow can be approximated as:
\vspace{-5pt}
\begin{equation}\label{eq:LinTrajModel}
    \bm{u}_n(\bx) \approx
    \textstyle \, \sum_{i=1}^R \, \bm{q}_i(n)\bm{v}_i(\bx) \,\, ,
\vspace{-5pt}
\end{equation}
where $\bm{v}_i(\bx)$ is the weight that needs to be applied on the $i$-th correspondence basis element, in order to get the correspondence vector for the point location $\bm{x}$. In other words, the shape flow for every point $\bm{x}$ is described as a linear combination of basis elements that is controlled by the coefficients $\bm{v}_i(\bm{x})$.
The values of the $i$-th coefficient for all the points $\bm{v}_i(\bm{x})$ can be interpreted as an image defined on $\Omega$. Using the correspondence basis, the determination of the shape flow boils down to the determination of the set of coefficients $\bm{v}_i(\bm{x})$. The above representation of shape flow, constrains the correspondence vectors to lie on a subspace and, therefore, acts as a low-rank prior that enforces coherency of the shape registration result over the whole training dataset of shapes.

@@

\begin{figure}[b!]
    \centering
    \includegraphics[width=0.9\columnwidth]{resources/models}
    \caption{Dense shape models build for faces and ears. Dense shapes are presented as grid for better visualisation.}
    \label{fig:dense_models}
\end{figure}

% ---------------------------------------------------------------------------------------------------------------------------------------------------
\vspace{0.3cm}
{\refstepcounter{steps}\label{sec:step3}\subsection*{Step 3: Shape Flow Estimation}}
% \subsection{Shape Flow Estimation}
% \label{sec:step3}

@@

Let $\bm{d}(\bx;n)$, $\bm{d}(\bx;0):\Omega\rightarrow \R^{N_c}$ be the $n$-th training SVS image and the reference SVS image respectively. Following \cite{Garg:2013hu}, we propose to estimate the shape flow over all training images by minimizing the following energy:
\vspace{-5pt}
\begin{align}
E_{sf} & =\alpha
\int_{\Omega}\sum_{n=1}^{N_t} \|\bm{d}(\bx+\bm{u}_n(x);n)-\bm{d}(x;0)\| \ud \bx \label{eq:costfunc}\\
    &+ \beta \int_{\Omega}\sum_{n=1}^{N_t}\|\bm{u}_n(\bx)-\sum_{i=1}^R\bm{q}_i(n)\bm{v}_i(\bx)\|^2 \ud \bx \label{eq:lowrank}\\
    &+
\int_\Omega  \sum_{i=1}^R \,\, \left \|    \nabla \bm{v}_i(\bx)    \right \|  \,\ud \bx \label{eq:TVterm}
\vspace{-5pt}
\end{align}
This energy consists of two sets of unknown shape flows that are relatively close to each other: (i) $\bm{u}_n(x)$ which tries to explain the data from the input SVS images, and (ii) the shape flow determined by the correspondence basis coefficients $\bm{v}_i(\bx)$ that are spatially regularised and enforce a low-rank prior.
%We minimize this energy jointly with respect to $\bm{u}_n(\bx)$ and $\bm{v}_i(\bx)$.
%The positive parameters $\alpha$ and $\beta$ weigh the balance between the terms of the energy.

@@

\vspace{0.3cm}
{\refstepcounter{steps}\label{sec:step4}\subsection*{Step 4: Dense and Patch-Based Deformable Models}}
% \subsection{Dense and Patch-Based Active Appearance Models}

### 2016-04-07 contents/introduction.tex
minor changes and acknowledgments

\item The manual annotation of such databases that has been undertaken by several research teams \cite{sagonas_iccv_300w_2013,sagonas2016faces,charles2013domain,dantone2014body,andriluka14cvpr}.

@@

\item Annotating with regards to consistent landmarks is an extremely time consuming, tedious and labour intensive work \cite{sagonas_iccv_300w_2013}, which is usually performed by a trained person. Furthermore, for various object classes, it requires a highly skilled person in order to identify and annotate landmarks in a consistent manner. For example, the human ear has very complicated inner structures (helix, crus antihelicis, scapha, tragus, lobe etc.) which remarkably differ between different ears. Moreover, certain ear parts, such as fossa triangularies and crus helicis, do not appear in all ears and their visibility is highly sensitive to the head pose and illumination variation. Another such example is the human body, which is generally annotated with regards to a number of landmarks that intuitively correspond to a set of body joints. For most body pose databases, annotation task was undertaken by a crowd-sourcing Internet marketplace, so-called Amazon Mechanical Turk (AMT). Unfortunately, this resulted in acquiring inconsistent and inaccurate annotations, in many cases\footnote{In the case of faces, the quality of annotations produced from AMT are extremely inaccurate and cannot, by any means, be compared with the ones provided by the recent 300W competition \cite{sagonas_iccv_300w_2013}.} (please see Figure \ref{fig:wrong_anno}). As it was also recently pointed out \cite{tompson2015efficient}, the inconsistencies in body joint annotations may also render the comparison between different human pose estimation methodologies irrelevant.

@@

To illustrate how time consuming careful annotation of a complex deformable object is, we lay down our own experience based on the human ear. A trained annotator needs an average of 4 minutes per image for the manual annotation of 55 landmarks. This means that the annotation of 1000 images requires a total of about 67 hours. Furthermore, the quality of training as well as fatigue largely influence the annotation accuracy. Hence, a second pass on the annotated data is, in many cases, necessary. Due to the fact that manual annotation is a costly and labour-intensive procedure, unsupervised learning of deformable models for the task of object alignment has recently attracted some attention \cite{frey2003learning, baker2004automatic, cootes2004groupwise, jojic2006escaping, Huang2006, kokkinos2007unsupervised, jiang2009learning, liu2009simultaneous, Zhang2012}. However, because the problem of fully unsupervised discovery of the deformations of arbitrary objects is difficult and ill-posed, the limited number of methods that have been proposed for the task cannot be directly applied to arbitrary collection of ``in-the-wild" images. On the other hand, the method of \cite{antonakos2014automatic}, which can deal with ``in-the-wild" images, requires a set of consistently annotated sparse shapes to perform deformable face congealing.

@@

In this paper, we propose a solution for annotating an object with regards to its deformations that requires considerably less effort compared to manual annotation and, at the same time, can be used to define statistical deformations for objects without a consistent set of landmarks. We employ the proposed method in order to construct SDMs based on the outline of human body parts (i.e., arms and legs). While majority of proposed SDMs for human poses are trained on skeleton joints, training SDMs based on outline of human body parts have shown remarkable advantages~\cite{zuffi2012pictorial} (?). The proposed SDM can also be used to provide accurate and consistent annotations for several of the body joints (such as wrist, elbow etc.).
To this end, we argue and empirically demonstrate that it is better to annotate an object with regards to a set of continuous lines that describe its shape. An example is provided in Figure \ref{fig:intro}, which compares the standard landmark annotations that are employed in the current literature with the proposed curve annotations for arms, ears and faces. It becomes evident that the curve annotations avoid the inherent ambiguity of placing sparse landmarks and offer a richer description of the object's shape.

Furthermore, these curves can be automatically generated by recently proposed methods that perform discriminative segmentation of objects \cite{luo2013pedestrian,liu2015matching}.

@@

Additionally, we show that the proposed methodology can be applied on landmark localisation, even though it is not tailored for that task, achieving particularly good performance. Specifically, we explain how to build powerful dense SDMs that are suitable for objects that have rich interior texture but lack landmarks consistency. Furthermore, we show how to build a powerful patch-based SDMs on the sparse outline landmarks of objects that do not have semantically meaningful interior textures. Using the resulting outline patch-based SDM, we report state-of-the-art performance on the task of human body parts localisation on challenging databases. Finally, we show that the proposed patch-based SDM can be used to provide consistent annotations for different body parts.

@@

  \item We report state-of-the-art quantitative and qualitative results on the task of human body parts localisation by employing a patch-based SDM trained on the outline landmarks that are sampled by the dense correspondences. Our proposed model outperforms all current state-of-the-art techniques that are trained on skeleton joints.

>>>>

\item The manual annotation of such databases that has been undertaken by several research teams \cite{sagonas2016faces,charles2013domain,dantone2014body,andriluka14cvpr}.

@@

\item Annotating with regards to consistent landmarks is an extremely time consuming, tedious and labour intensive work \cite{sagonas_iccv_300w_2013}, which is usually performed by a trained person. Furthermore, for various object classes, it requires a highly skilled person in order to identify and annotate landmarks in a consistent manner. For example, the human ear has very complicated inner structures (helix, crus antihelicis, scapha, tragus, lobe etc.) which remarkably differ between different ears. Moreover, certain ear parts, such as fossa triangularies and crus helicis, do not appear in all ears and their visibility is highly sensitive to the head pose and illumination variation. Another such example is the human body, which is generally annotated with regards to a number of landmarks that intuitively correspond to a set of body joints. For most body pose databases, the annotation task was undertaken by a crowd-sourcing Internet marketplace, so-called Amazon Mechanical Turk (AMT). Unfortunately, this resulted in acquiring inconsistent and inaccurate annotations, in many cases\footnote{In the case of faces, the quality of annotations produced from AMT are extremely inaccurate and cannot, by any means, be compared with the ones provided by the recent 300W competition \cite{sagonas_iccv_300w_2013,sagonas2016faces}.} (please see Figure \ref{fig:wrong_anno}). As it was also recently pointed out \cite{tompson2015efficient}, the inconsistencies in body joint annotations may also render the comparison between different human pose estimation methodologies irrelevant.

@@

To illustrate how time consuming careful annotation of a complex deformable object is, we lay down our own experience based on the human ear. A trained annotator needs an average of 4 minutes per image for the manual annotation of 55 landmarks. This means that the annotation of 1000 images requires a total of about 67 hours. Furthermore, the quality of training as well as fatigue greatly influence the annotation accuracy. Hence, a second pass on the annotated data is, in many cases, necessary. Due to the fact that manual annotation is a costly and labour-intensive procedure, unsupervised learning of deformable models for the task of object alignment has recently attracted some attention \cite{frey2003learning, baker2004automatic, cootes2004groupwise, jojic2006escaping, Huang2006, kokkinos2007unsupervised, jiang2009learning, liu2009simultaneous, Zhang2012}. However, because the problem of fully unsupervised discovery of the deformations of arbitrary objects is difficult and ill-posed, the limited number of methods that have been proposed for the task cannot be directly applied to arbitrary collection of ``in-the-wild" images. On the other hand, the method of \cite{antonakos2014automatic}, which can deal with ``in-the-wild" images, requires a set of consistently annotated sparse shapes to perform deformable face congealing.

@@

In this paper, we propose a solution for annotating an object with regards to its deformations that requires considerably less effort compared to manual annotation and, at the same time, can be used to define statistical deformations for objects without a consistent set of landmarks. We employ the proposed method in order to construct SDMs based on the outline of human body parts (i.e., arms and legs). The proposed SDM can also be used to provide accurate and consistent annotations for several of the body joints (such as wrist, elbow etc.).
To this end, we argue and empirically demonstrate that it is better to annotate an object with regards to a set of continuous lines that describe its shape. An example is provided in Figure \ref{fig:intro}, which compares the standard landmark annotations that are employed in the current literature with the proposed curve annotations for arms, ears and faces. It becomes evident that the curve annotations avoid the inherent ambiguity of placing sparse landmarks and offer a richer description of the object's shape.
%
Furthermore, these curves can be automatically generated by recently proposed methods that perform discriminative segmentation of objects \cite{luo2013pedestrian,liu2015matching}.
%
Note that the work in~\cite{zuffi2012pictorial} is the only one that shows that training SDMs based on the outline contours of the human body parts has considerable advantages compared to using the sparse skeleton joints, as done by the majority of existing SDMs for human pose.

@@

Additionally, we show that the proposed methodology can be applied on landmark localisation, even though it is not tailored for that task, achieving particularly good performance. Specifically, we explain how to build powerful dense SDMs that are suitable for objects that have rich interior texture but lack landmarks consistency. Furthermore, we show how to build a powerful patch-based SDM on the sparse outline landmarks of objects that do not have semantically meaningful interior textures. Using the resulting outline patch-based SDM, we report state-of-the-art performance on the task of human body parts localisation on challenging databases. Finally, we show that the proposed patch-based SDM can be used to provide consistent annotations for different body parts.

@@

  \item We report state-of-the-art quantitative and qualitative results on human body parts localisation by employing a patch-based SDM trained on the outline landmarks that are sampled by the dense correspondences. Our proposed model outperforms all current state-of-the-art techniques that are trained on skeleton joints.
